---
title: "LTE Exploration"
author: "Ahmed Tarek Sayed"
date: "June 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

importing Libraries
```{r}
library(tidyverse)
library(Amelia)
library(readxl)
library(PerformanceAnalytics)
library(ggcorrplot)
```
Reading updated version
```{r}
X4G <- read_csv("E:/Academic/ITI/PROJECT/project network/lte/Data/DataSet/4G dataset updated(KPI Analysis Result).csv", 
    col_types = cols(`Cell FDD TDD Indication` = col_skip(), 
        Date = col_datetime(format = "%m/%d/%Y %H:%M"), 
        Integrity = col_skip(), L.ChMeas.RI.Rank3 = col_integer(), `(HU) Average UE Number` = col_double(),
        L.ChMeas.RI.Rank4 = col_integer(), 
        `LocalCell Id` = col_skip(), `eNodeB Function Name` = col_skip(), 
        `eNodeB Name` = col_skip()))
glimpse(X4G)

LTE <- X4G
```


Importing Data Set

```{r}
#LTE <- read_csv("E:/Academic/ITI/PROJECT/project network/lte/Data/Mod Data/LTE dataset (CSV).csv", 
 #   col_types = cols(Date = col_datetime(format = "%m/%d/%Y %H:%M")))
```
Checking Nulls

```{r}
missmap(LTE)
```

```{r}
LTE
```
Selecting and renamming Variables
```{r}
#LTE <- LTE %>% select(Date, `Cell Name`,`(HU) Average CQI Per Cell`, `(HU) Average UE Number`, `(HU) User DL Average Throughput (Mbps)`, `(HU) DL Traffic Volume (MB)`, `(HU) UL Traffic Volume (MB)`, `(HU) User UL Average Throughput (Mbps)`, `(HU) DL PRB Utilization(%)`, starts_with('L.RA'))

LTE <- LTE %>% rename(CQI = `(HU) Average CQI Per Cell`,avg_ue = `(HU) Average UE Number`,avg_dl_cell_throughput = `(HU) Cell DL Average Throughput (Mbps)`,avg_dl_user_throughput = `(HU) User DL Average Throughput (Mbps)`,avg_ul_user_throughput = `(HU) User UL Average Throughput (Mbps)`,dl_traffic = `(HU) DL Traffic Volume (MB)`,ul_traffic = `(HU) UL Traffic Volume (MB)`,ul_traffic = `(HU) UL Traffic Volume (MB)`,avg_ul_user_throughput = `(HU) User UL Average Throughput (Mbps)`,prb_utilization = `(HU) DL PRB Utilization(%)`, avg_mcs = `(HU)Avg MCS`)

LTE
```



Checking the other variables where CQI is null
```{r}
LTE %>% filter(is.na(CQI))  %>% summary()
```
How many are they ?

```{r}
LTE %>% filter(is.na(CQI)) %>% summarise(n())
```

All vars are either 0 or NA where CQI is null except for Prb which shows a very low values (Clearly beacaus it is not utilized)
How about when there is no nulls
```{r}

LTE %>% filter(!is.na(CQI))  %>% summary()
```
Clearly the prb has increased once the CQI is not null (the cell is active) and other variables carry a representable number
Checking if ever variables carry 0 value similtinously (Which may indicates also a non functioning cell)
```{r}
LTE %>% filter(CQI ==0,avg_ue == 0, avg_dl_user_throughput == 0 ) %>% summarise(n())
```

Now I can remove nulls confidently that they have no value and no other condition may indicate a non functioning cell that can corrupt our model

```{r}
LTE <- LTE %>% filter(!is.na(CQI))
```
Lets see how many rows remaining
```{r}
length(LTE$CQI)
```
Convient amount representing 99.999% of the data

```{r}
summary(LTE)
```



Lets look at our target variable

from the summary we can see that
it is a little right skewed means that there is users in some cells gets a extremly high avg throughput
OK lets look
```{r}
LTE %>% ggplot(aes(x = avg_dl_user_throughput)) + geom_histogram(aes(y = ..density.., fill = ..count..)) + stat_function(fun = dnorm,color = 'red' ,args = list(mean = mean(LTE$avg_dl_user_throughput, na.rm = T),sd= sd(LTE$avg_dl_user_throughput, na.rm = T)))
```
Using log the variable
```{r}
LTE %>% ggplot(aes(x = log(avg_dl_user_throughput))) + geom_histogram(aes(y = ..density.., fill = ..count..)) + stat_function(fun = dnorm,color = 'red' ,args = list(mean = log(mean(LTE$avg_dl_user_throughput, na.rm = T))),sd= sd(log(LTE$avg_dl_user_throughput), na.rm = T))
```
Note to self:-
(I have a question regarding the Avg ue, What is the granular time windiw that is avereged for each hour ?)
unpivot and see what is the most used distance from indexes
use standard deviasio as a method of detcting abnormalty (after applying significant groupping)

#Now Lets see what are those extreme cells ? are the distribution of avg user throughput related to certain mode of transmission and reception or even the band ?
importing and joining the LTE join the data sets to see
```{r}
LTE_cells <- read_excel("E:/Academic/ITI/PROJECT/project network/lte/Data/DataSet/LTE cells.xlsx")
LTE_cells <- LTE_cells %>% rename(`Cell Name` = `*Cell Name`)

LTE_det <- left_join(LTE, LTE_cells)
LTE_det <-  LTE_det %>% rename(band = `*Downlink bandwidth`, trans_recep_mode = `*Cell transmission and reception mode`)
```
Making sure of complete join
```{r}
LTE_det %>%  filter(is.na(band) & is.na(trans_recep_mode))%>% summarise(n_distinct(`Cell Name`), n())
```
There is 226 cell without a refrence in LTE_cells reprsenting 4225 instance

#Now seeing if the Cell band affects the throughput
```{r}
LTE_det %>% group_by(band) %>% summarise(mean(avg_dl_user_throughput, na.rm = T), n_distinct(`Cell Name`), sd(avg_dl_user_throughput, na.rm = T))
```
Notice that the N50 band has the highest standard deviasion in all bands
```{r}
LTE_det %>% ggplot(aes(x = log(avg_dl_user_throughput))) + geom_histogram(aes(fill = LTE_det$band))
```



```{r}
LTE_det %>% ggplot(aes(x= band, y = avg_dl_user_throughput)) + geom_boxplot()
```

making annova to see diffrence

```{r}
LTE_fin <- LTE_det %>% filter(is.finite(log(avg_dl_user_throughput)))
aov_band <- aov(data = LTE_fin, log(LTE_fin$avg_dl_user_throughput) ~ LTE_fin$band)

summary(aov_band)
```

```{r}
TukeyHSD(aov_band)
```
Checjing homogenicity  of variance
```{r}
plot(aov_band,1)
plot(aov_band,2)

```

We can now conclode that :-
All bands are significantly diffrent from each other except for N75 and N50 which can be interpreted as one

Does the mode affets with the band

```{r}

LTE_det %>% ggplot(aes(x= trans_recep_mode, y = avg_dl_user_throughput)) + geom_boxplot()
```
The  N75 band has only one trans mode (1T1R) which may explain why N75 is not diffrent from N50 despite the higher technology used

Now seeing if the trans mode make differ

```{r}
aov_trans <- aov(data = LTE_fin, log(LTE_fin$avg_dl_user_throughput) ~ LTE_fin$trans_recep_mode)

summary(aov_trans)
```


```{r}
TukeyHSD(aov_trans)
```
We can see that 4T4R doesnot differ from 2T4R and 1T2R is lower than 1T1R

Lets see both the band and the trans mode
```{r}
LTE_det %>% ggplot(aes(y= avg_dl_user_throughput, x = band)) + geom_boxplot(aes(color = trans_recep_mode))
```

```{r}
LTE_det %>% group_by(band, trans_recep_mode) %>%  summarise(mean(avg_dl_user_throughput, na.rm = T), n_distinct(`Cell Name`), sd(avg_dl_user_throughput, na.rm = T))
```
##We can see that the increased sd is Mainly band related not a trans mode related


Joining the region table to see if the region or subregion affects

```{r}
region <- read_excel("E:/Academic/ITI/PROJECT/project network/lte/Data/DataSet/sub-region.xlsx", 
    col_types = c("text", "blank", "blank", 
        "blank", "blank", "blank", "text", 
        "text", "text", "blank"))
LTE_det<-  LTE_det %>% select(-`*eNodeB Name`, -`Cell active state`, -L.RA.TA.UE.Index10, -L.RA.TA.UE.Index11, -L.RA.TA.UE.Index11,-L.RA.TA.UE.Index8, -L.RA.TA.UE.Index9)
region <- region %>% rename(`Cell Name` = CELL)
LTE_det <- left_join(LTE_det, region )
```
Note to self:-
Does the indeces and no of ue rsemblant?
Does the the indices whn weighted is more correlated to throughput ?

Now seeing the un joined
```{r}
LTE_det %>% filter(is.na(Region)) %>% summarise(no_region_instances = n(), no_region_cells = n_distinct(`Cell Name`))
```

```{r}
missmap(LTE_det)
```
Now all the missing instances represents 1% of the data

Now lets see if the throghuput differ from region to another

```{r}
LTE_det %>% group_by(Region) %>% summarise(mean(avg_dl_user_throughput, na.rm = T), n_distinct(`Cell Name`), sd(avg_dl_user_throughput, na.rm = T))
```

```{r}
LTE_det %>% ggplot(aes(Region,avg_dl_user_throughput)) + geom_boxplot()
```

a diffrence is clear but is it the region or just the distribution of technologies across the regions ?
```{r}
LTE_det %>% ggplot(aes(Region,avg_dl_user_throughput)) +geom_boxplot(aes(color = LTE_det$band ))
```
 
It looks like regardless of the band the diffrent regions has a diffrent throughput

Lets look at the ue traffic in the region if it what makes the diffrence ?
```{r}
LTE_det %>% ggplot(aes(Region,avg_ue)) + geom_boxplot() + ylim(NA, 100)
```
 Looks like the avg UE per cell does not differ across regions except for alexandriawhich shows a much lower avg ue
 
```{r}
LTE_det %>% ggplot(aes(Region,LTE_det$dl_traffic)) +geom_boxplot() + ylim(NA,10000 )
```
 What I can conclude is that Alexandria has the lowest avg traffic and UE number which can contribute to its high throughput
 
 Delta is showing a strange behaviour that regardless of the relatively low traffic and avg UE the throughput is low, Which may be caused by using only N25 band in delta and no other technology
 
 Testing regions diffrence
```{r}
aov_reg <- aov(data = LTE_det, LTE_det$avg_dl_user_throughput ~ LTE_det$Region*LTE_det$band)

summary(aov_reg)
```
 
```{r}
TukeyHSD(aov_reg)
```
 There is a statistical diffrence across all regions except for the upper-egypt and canal
 
 Discarding Upload link related variables
```{r}
LTE_det <- LTE_det %>% select(-avg_dl_cell_throughput, -ul_traffic, -avg_ul_user_throughput)
```
 
 Now lets look at the numerical variables and how they contribute to  the throughput and UE and Traffic
 
```{r}
tokeep <- which(sapply(LTE_det,is.numeric))
corr <- round(cor(LTE_det[complete.cases(LTE_det),tokeep]),1)
ggcorrplot(corr, hc.order = TRUE,  lab = TRUE)
```
Removing all nulls from the data set

```{r}
LTE_det <- LTE_det[complete.cases(LTE_det),]
```


```{r}
cor(LTE_det$avg_dl_user_throughput, LTE_det$dl_traffic)
```
```{r}
cor(LTE_det$avg_dl_user_throughput, LTE_det$avg_ue)
```

avg UE has a correlation of -0.3 to throughput and avg traffic is -0.1 which is very low correlation


Can we derive a new variable that is the the traffic per user Will it be more relvant 
```{r}
LTE_det$traf_user_rat = LTE_det$dl_traffic / LTE_det$avg_ue
cor(LTE_det$avg_dl_user_throughput, LTE_det$traf_user_rat)
LTE_det <- LTE_det %>% select(-traf_user_rat)
```
Shown a lower correlation to our throughput variable than the avg_ue

How about weighting the indexes and see if the weighted sum is more correlated than each and avg ue to both traffic and throughput
```{r}
LTE_det$wted_index = LTE_det$L.RA.TA.UE.Index0*78 + LTE_det$L.RA.TA.UE.Index1*273 + LTE_det$L.RA.TA.UE.Index2 * 390 + LTE_det$L.RA.TA.UE.Index3 * 603 + LTE_det$L.RA.TA.UE.Index4 *1532 + LTE_det$L.RA.TA.UE.Index5 * 2730 + LTE_det$L.RA.TA.UE.Index6 * 5070 + LTE_det$L.RA.TA.UE.Index7 * 10530
```

```{r}
tokeep <- which(sapply(LTE_det,is.numeric))
corr <- round(cor(LTE_det[complete.cases(LTE_det),tokeep]),1)
ggcorrplot(corr, hc.order = TRUE,  lab = TRUE)
```

 